---
title: "Figure 1 doublets"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true

---

```{r}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(ROCit)
  library(ggplot2)
  library(cowplot)
})
devtools::load_all("~/bioinfo/robinson/scDblFinder/")
theme_set(theme_cowplot(font_size = 12))

# we load the dataset
lf <- list.files("../datasets/", pattern="SCE", full.names=TRUE)
names(lf) <- sapply(strsplit(basename(lf),".",fixed=TRUE), head, 1)
CDs <- lapply(lf, FUN=function(x){ as.data.frame(colData(readRDS(x))) })
```

```{r}
res <- dblres <- readRDS("../data/doublets.rds")
scoresNames <- c("scran.score"="scran", "scDblFinder"="scDblFinder", "DF.score"="DoubletFinder", "hybrid_score"="scds")
cols <- c("#4477AA", "#DDCC77", "#117733", "#CC6677")
names(cols) <- as.character(scoresNames)

pw <- function(i, dot.size=4, ...){
  scores <- res[[i]][,intersect(names(scoresNames),colnames(res[[i]]))]
  colnames(scores) <- scoresNames[colnames(scores)]
  plotROCs(scores, factor(res[[i]]$demuxlet_cls, c("SNG","DBL")), dot.size=dot.size) +
    ylim(0,1) + scale_color_manual(values=cols)
}

p1 <- pw(1) + theme(legend.position="none")
p2 <- pw(2) + theme(legend.position="none")
p3 <- pw(3) + theme(legend.position=c(0.25,0.25))
```

```{r}
rt <- data.frame(
  method = structure(c(3L, 4L, 2L, 1L, 3L, 4L, 2L, 1L, 
3L, 4L, 2L, 1L), .Label = c("DoubletFinder", "scDblFinder", "scds", 
"scran"), class = "factor"),
  dataset = structure(c(1L, 1L, 1L, 
1L, 2L, 2L, 2L, 2L, 3L, 3L, 3L, 3L), .Label = c("mixology10x3cl", 
"mixology10x5cl", "demuxlet"), class = "factor"),
  time = c(14.52, 26.36, 7.82400000000007, NA, 42.74, 
250.14, 21.8199999999999, 678.81, 100.05, 823.09, 54.0140000000001, 
956.63)
)

pt <- ggplot(rt, aes(method, time, fill=method)) + geom_col() + facet_wrap(~dataset, scale="free_x") + coord_flip() + xlab("") + ylab("Running time (s)") + scale_fill_manual(values=cols) + theme(legend.position="none")

# le <- get_legend(plotROCs(scores, factor(CDs[[i]]$demuxlet_cls, c("SNG","DBL"))))
```


```{r}
mis <- readRDS("../data/misclassification.rds")
lf <- list.files("../datasets/", full.names = TRUE)
lf <- lf[grep("simMix",lf,invert=TRUE)]
names(lf) <- sapply(strsplit(basename(lf),"\\."), FUN=function(x) x[1])
CDs <- lapply(lf, FUN=function(x) colData(readRDS(x)))
dbl <- lapply(CDs, FUN=function(x) x[,c("pldf.ratio","pldf.class")])
dbl <- dbl[intersect(names(dbl),names(mis))]
for(n in names(dbl)){
  dbl[[n]]$pct_misclassified_cor <- mis[[n]]$corrected_pct_mis[row.names(dbl[[n]])]
  dbl[[n]]$pct_misclassified <- mis[[n]]$pct_misclassified[row.names(dbl[[n]])]
}
d <- as.data.frame(data.table::rbindlist(lapply(dbl,as.data.frame), idcol="dataset"))
d$pldf.class <- factor(d$pldf.class, c("doublet","singlet"))
levels(d$pldf.class) <- c("DBL","SNG")
p4 <- ggplot(d, aes(pldf.class, pct_misclassified, fill=pldf.class)) + geom_violin() + facet_wrap(~dataset, nrow = 1) + theme(axis.title.x=element_blank(), legend.position="none") + ylab("% misclassified") + stat_summary(fun.y="mean", geom="point")
```



```{r figure1, fig.height=6, fig.width=7}
p1m <- plot_grid(p1, p2, p3, nrow=1)
plot_grid(p1m, pt, p4, rel_heights=c(3.5,2,2), labels = LETTERS[1:3], ncol = 1, hjust=0)
```

**A:** Receiver operating characteristic (ROC) curves of the tested doublet detection methods for three datasets with SNP-identified doublets. **B:** Running time of the different methods (DoubletFinder failed on one of the datasets). **C:** Rate of misclassification of the cells identified by scDblFinder as doublets (DBL), across a large range of clustering analysis. Even in datasets which should not have neotypic doublets, the cells identified as such tend to be misclassified.
