---
title: "Supplementary Figures relative to dimensionality reduction"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r, include=FALSE}
if(!exists('FIG_NB')){
  FIG_NB <- 0
  getFigNb <- function(increment=FALSE){
    if(increment) FIG_NB <<- FIG_NB + 1
    FIG_NB
  }
}
```

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(reshape2)
  library(dplyr)
  library(pipeComp)
  library(viridis)
})
theme_set(theme_cowplot(font_size = 13))
source("../misc_functions.R")
```

# Supplementary Figure `r getFigNb(TRUE)`

```{r dimred, fig.width=10.5, fig.height=6}
res <- readRDS("../data/results_dimred2.rds")
H <- renameHrows(
  scrna_evalPlot_silh( res, reorder_rows=TRUE, 
                       heatmap_legend_param=list(direction="horizontal")),
  function(x) gsub("norm\\.","",x))
p1 <- grid.grabExpr(draw(H, heatmap_legend_side="bottom"))

p2 <- renameHrows(
  evalHeatmap( res, step="clustering", what=c("MI","min_pr"), agg.by=c("norm","dr"), 
               reorder_rows=TRUE, anno_legend=FALSE, title=c("MI","Minimum precision")) + 
  evalHeatmap(res, step="clustering", what=c("ARI"), agg.by=c("norm","dr"), 
              filterExpr=n_clus==true.nbClusts, anno_legend=FALSE, title="ARI at true k"),
  function(x) gsub("norm\\.","",x)
)
p2 <- grid.grabExpr(draw(p2))

plot_grid(p1,p2, nrow=2, labels=LETTERS[1:2])
```

### Supplementary Figure `r getFigNb()`
**Evaluation of common dimensionality reduction methods. A:** Minimum (left) and average (right) silhouette width per subpopulation resulting from combinations of normalization and dimension reductions. **B:** Clustering accuracy, measured by mutual information (MI), minimum subpopulation precision, and adjusted Rand index (ARI) at the true number of clusters.

\newpage

# Supplementary Figure `r getFigNb(TRUE)`


```{r, fig.width=8, fig.height=8.5}
res <- readRDS("../data/clust_endSummary.rds")
res2 <- res[which(res$clust=="clust.seurat" & res$k==20 & res$dims>2),]
res2 <- res2[which(res2$dr=="seurat.pca" & res2$dataset!="Kumar"),]
res2$diff.nbClusters <- res2$nbClusters-truek[res2$dataset]
res3 <- aggregate(res2[,"diff.nbClusters"], by=res2[,c("dataset","dims","resolution")], FUN=mean)
colnames(res3)[4] <- "nb"
ll <- lapply(split(res3, res3$dataset), FUN=function(x){
  ggplot(x, aes(factor(dims), factor(resolution), fill=nb)) + stat_bin2d() + 
    scale_fill_viridis() + ylab("Resolution") + xlab("Number of dimensions") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
})
for(i in names(ll)) ll[[i]] <- ll[[i]] + ggtitle(i)
plot_grid(plotlist = ll, nrow=3)
```

### Supplementary Figure `r getFigNb()`
Mean difference between the number of detected clusters and the number of real subpopulations, depending on the resolution and number of dimensions used. Based on sctransform and seurat PCA. Increasing the number of dimensions tends to decrease the number of identified clusters, especially at resolutions around the default value.

\newpage

# Supplementary Figure `r getFigNb(TRUE)`

```{r, fig.width=8, fig.height=8.5}
res3 <- aggregate(res2[,"ARI",drop=FALSE], by=res2[,c("dataset","dims","resolution")], FUN=mean)
ll <- lapply(split(res3, res3$dataset), FUN=function(x){
  ggplot(x, aes(factor(dims), factor(resolution), fill=ARI)) + stat_bin2d() + 
    scale_fill_viridis() + ylab("Resolution") + xlab("Number of dimensions") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
})
for(i in names(ll)) ll[[i]] <- ll[[i]] + ggtitle(i)
plot_grid(plotlist = ll, nrow=3)
```

### Supplementary Figure `r getFigNb()`

Adjusted Rand Index of clustering depending on the resolution and number of dimensions used. Based on sctransform
and seurat PCA.

\newpage

# Supplementary Figure `r getFigNb(TRUE)`


```{r, fig.width=7.5, fig.height=8.5}
cl <- readRDS("../data/results_dimensionality.rds")$evaluation$clustering
cl$delta.nbClust <- cl$n_clus-cl$true.nbClusts

cl$dims <- gsub("maxLikGlobal$","maxLikGlobal(k=20)",cl$dims)
cl$dims <- gsub("maxLikGlobal10$","maxLikGlobal(k=10)",cl$dims)
cl <- cl[cl$norm=="norm.seurat" & !(cl$dims %in% c("10","30","essLocal.b")),]

ggplot(cl, aes(x=dims, y=delta.nbClust, fill=dims)) + geom_violin() + 
  facet_wrap(~dataset, scales = "free_x", ncol = 2) + coord_flip() + 
  stat_summary(size=0.25, fun.data = mean_se, na.rm = TRUE) + 
  geom_hline(yintercept = 0, linetype="dashed") + xlab("") +
  theme(legend.position="none", text=element_text(size=10)) + 
  ylab("Deviation from true number of clusters")
```

### Supplementary Figure `r getFigNb()`

Deviation from the true number of clusters using different number of principal components (based the indicated dimensionality estimates) of the same Seurat-based PCA. Default pipeline parameters were used for the other steps, and the distributions represent the different resolutions of Seurat clustering. Across datasets, MCVR and maxLikGlobal appear to depart less from the true number of clusters.

\newpage

# Supplementary Figure `r getFigNb(TRUE)`

```{r, include=FALSE, eval=FALSE}
source(system.file("extdata", "scrna_alternatives.R", package="pipeComp"))
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(Seurat)
  library(scater)
  library(intrinsicDimension)
})
datasets <- list.files("../datasets", full.names = TRUE)
names(datasets) <- gsub("\\..*", "", basename(datasets))
datasets <- lapply(datasets, readRDS)
datasets <- lapply(datasets, filt.mad)
datasets <- lapply(datasets, norm.seurat)
datasets <- lapply(datasets, sel.deviance)

met <- c("seurat.pca", "scran.denoisePCA", "scran.runPCA")
res <- list()
for(i in met) {
  res[[i]] <- list()
  print(i)
  a <- lapply(datasets, function(x) {
    switch(i,
           seurat.pca = seurat.pca(x),
           scran.denoisePCA = scran.denoisePCA(x),
           scran.runPCA = scran.runPCA(x))
  })
  x <- lapply(a, function(x) reducedDim(x, "PCA"))
  rm(a); gc()
  ks <- seq(5, 40, 5)
  for (d in names(datasets)) {
    res[[i]][[d]]$ndims <- c()
    res[[i]][[d]]$k <- c()
  }
  for (k in ks) {
    print(paste("k:", k))
    for (d in names(datasets)) {
      res[[i]][[d]]$ndims <- c(res[[i]][[d]]$ndims,
                               as.numeric(maxLikGlobalDimEst(x[[d]], k=k, unbiased=TRUE)))
      res[[i]][[d]]$k <- c(res[[i]][[d]]$k, k)
    }
  }
}
saveRDS(res, "../data/maxLikGlobal.rds")
```

```{r, fig.width=7, fig.height=5}
res <- readRDS("../data/maxLikGlobal.rds")
res <- dplyr::bind_rows(lapply(res$seurat.pca, as.data.frame), .id="dataset")
cols <- pipeComp::getQualitativePalette(9)
names(cols) <- unique(res$dataset)
ggplot(res, aes(k, ndims, colour=dataset)) + geom_point() + geom_line() + 
  scale_colour_manual(values=cols) + xlab("Number of nearest neighbors (k)") +
  ylab("Estimated dimensionality") + 
  ggtitle("intrinsicDimension::maxLikGlobalDimEst")
```

### Supplementary Figure `r getFigNb()`
Estimates of dimensionality by the `intrinsicDimension::maxLikGlobalDimEst` method using various 'reasonable' numbers of nearest neighbors (`k` parameter).

