---
title: "Supplementary Figures relative to feature selection"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r, include=FALSE}
if(!exists('FIG_NB')){
  FIG_NB <- 0
  getFigNb <- function(increment=FALSE){
    if(increment) FIG_NB <<- FIG_NB + 1
    FIG_NB
  }
}

suppressPackageStartupMessages({
  library(ggplot2)
  library(ggrepel)
  library(cowplot)
  library(ComplexHeatmap)
  library(pipeComp)
})
theme_set(theme_cowplot(font_size = 11))
source("../misc_functions.R")
```

# Fig S`r getFigNb(TRUE)`

```{r, fig.align="center", echo=FALSE, out.height="0.95\\textheight"}
knitr::include_graphics("../main_figures/overall.pdf")
```

### Fig S`r getFigNb()`

**Overview of the combination of main alternative parameters.** Unsupervised clustering of the results (above) and top results (below). The color-mapping schemes are the same as described in the main figures.

\newpage

# Fig S`r getFigNb(TRUE)`


```{r, fig.width=6, fig.height=4}
ov <- readRDS("../data/overall.rds")
cl <- ov$evaluation$clustering
cl$norm <- relevel(cl$norm, "norm.seurat")
cl$doubletmethod <- relevel(cl$doubletmethod, "none")
cl$filt <- relevel(cl$filt, "none")
cl$sel <- relevel(cl$sel, "sel.vst")
cl$clustmethod <- relevel(cl$clustmethod, "clust.seurat")

# checking for interactions:
mod <- lm(ARI~resolution*dataset+doubletmethod*sel*filt*norm*clustmethod*dims, data=cl)
co <- as.data.frame(coef(summary(mod)))
coi <- co[grep(":",row.names(co)),]
coi <- coi[grep("dataset", row.names(coi), invert=TRUE),]
colnames(coi) <- c("Estimate","SE","t","p.value")
coi <- coi[order(coi$p.value),]
coi$FDR <- p.adjust(coi$p.value, method="fdr")
row.names(coi) <- gsub("selsel","sel",gsub("normnorm","norm",row.names(coi)))
coi$interaction <- gsub(":"," +\n", row.names(coi))
coi2 <- coi[coi$FDR < 0.05,]
ggplot(coi, aes(Estimate, -log10(p.value), label=interaction)) + geom_point() + 
  geom_point(data=coi2, colour="red") + geom_label_repel(data=coi2) + xlab("ARI coefficient") +
  ggtitle("Significant interactions between parameter values")
```

### Fig S`r getFigNb()`

Estimated coefficient and -log10 significance of all interaction coefficients in the linear model defined by 

`ARI~dataset*resolution+doubletmethod*sel*filt*norm*clustmethod*dims`

Those terms significant at a FDR < 0.05 are highlighted. The top interaction terms are:

```{r}
knitr::kable(coi[1:10,c(1,2,4,5)], digits=3)
```

\newpage

# Fig S`r getFigNb(TRUE)`

```{r}
rnfn <- function(x){
  x <- gsub("norm\\.","",x)
  x <- gsub("^doublet\\.","",x)
  x <- gsub("> none > ","> no filter > ",x)
  gsub("^none > ","",x)
}

H3 <- renameHrows( scrna_evalPlot_silh(ov, agg.by = c("doubletmethod","sel","norm"), row_split = rnfn(norm),
                                       heatmap_legend_param=list(direction="horizontal"),
                                       font_factor = 0.8, 
                                       row_names_max_width=unit(7, "cm")), rnfn)
H4 <- renameHrows(
     evalHeatmap(ov, step="clustering", what=c("MI","min_pr"), anno_legend=FALSE, title=c("MI","Min precision"),
                 agg.by=c("doubletmethod","sel","norm"), row_split=rnfn(norm), 
                 font_factor = 0.8, row_names_max_width=unit(7, "cm")) + 
    evalHeatmap(ov, step="clustering", what=c("ARI"), agg.by=c("doubletmethod","sel","norm"), 
                row_split=rnfn(norm),filter=n_clus==true.nbClusts, title="ARI at true k", 
                anno_legend=FALSE, font_factor = 0.8, row_names_max_width=unit(7, "cm")),
  rnfn)
```


```{r, fig.width=9, fig.height=8.5}
plot_grid( grid.grabExpr(draw(H3, heatmap_legend_side="bottom")), 
           grid.grabExpr(draw(H4)), nrow=2, labels=LETTERS[1:2])
```

### Fig S`r getFigNb()`

Silhouette widths of the real subpopulations (**A**) and clustering accuracy (**B**) using different combinations of methods.
